---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)


```

# mvabund

<!-- badges: start -->
[![](https://img.shields.io/badge/doi-10.1111/j.2041--210X.2012.00190.x-blue.svg)](https://doi.org/10.1111/j.2041-210X.2012.00190.x)
[![License](http://img.shields.io/badge/license-LGPL%20%28%3E=%202.1%29-brightgreen.svg?style=flat)](http://www.gnu.org/licenses/gpl-2.0.html) 
[![CRAN](http://www.r-pkg.org/badges/version/mvabund)](https://CRAN.R-project.org/package=mvabund) 
[![Downloads](http://cranlogs.r-pkg.org/badges/mvabund?color=brightgreen)](https://www.r-pkg.org/pkg/mvabund)
<!-- badges: end -->

The goal of mvabund is to .

### Installation

`mvabund` is available on [CRAN](https://CRAN.R-project.org/package=mvabund) and can be installed directly in R: 

```{r, eval = FALSE}
install.packages("mvabund")

library(mvabund)
```

Alternatively, you can install the **development** version of `mvabund` from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("aliceyiwang/mvabund")

library(mvabund)
```

### Getting Started

We highly recommend you taking a good read of our vignette over at [XXX]() before launching into the `mvabund`

### Show `mvabund` your support

```{r}
citation("mvabund")
```

### Spot a bug? 

Thanks for finding the bug! We would appreciate it if you can pop over to our [Issues page](https://github.com/aliceyiwang/mvabund/issues) and describe how to reproduce the bug! 